---
title: "p8105_hw3_rk3445"
author: "Rosie Kwon"
output: github_document
---

```{r}
library(tidyverse)
library(patchwork)
```

# Problem 1 

```{r}
library(p8105.datasets)
data("instacart")
```

The `instacart` dataset provides item-level records of customer grocery orders from the Instacart delivery service. This dataset contains `r nrow(instacart)` observations and `r ncol(instacart)` variables (columns), where each row represents a single product within an order.

Key variables in the dataset are following:

```{r}
user = 
  instacart |> 
  janitor::clean_names() |> 
  group_by(user_id) |> 
  summarize(n_order = n()) |> 
  arrange(desc(n_order)) 
```

* `user_id` is used for identifying each customers in the dataset, there are `r length(unique(instacart$user_id))` distinct customers in the instacart, which enables to analyze behavior at the customer level. Both `r user$user_id[1]` and `r user$user_id[2]` ordered the most, with `r user$n_order[1]` orders each.


```{r}
pop_product = 
  instacart |> 
  janitor::clean_names() |> 
  group_by(product_name) |> 
  summarize(n_product = n()) |> 
  arrange(desc(n_product)) 
```

* `product_name` and `product_id` are for identifying the specific product purchased, having `r length(unique(instacart$product_id))` unique products in the dataset. The most popular product are `r pop_product$product_name[1]` with `r pop_product$n_product[1]` orders , followed by `r pop_product$product_name[2]` with `r pop_product$n_product[2]` orders.


* `aisle_id`, `aisle`, `department_id` and `department` are variables for identifying the location of products and enabling grouping products. The products are organized into `r length(unique(instacart$aisle))` aisles and `r length(unique(instacart$department))` departments.


```{r}
day_order = 
  instacart |> 
  janitor::clean_names() |>
  mutate(
    order_dow_label = wday(order_dow + 1, label = TRUE, abbr = FALSE)
  ) |> 
  group_by(order_dow_label) |> 
  summarize(n_order = n())
```

* `order_dow`, `order_hour_of_day` indicate the time of orders from users. `days_since_prior_order` shows the how many days passed between orders with an average of `r round(mean(instacart$days_since_prior_order, na.rm = TRUE), 1)` days. Those variables are used to see shopping habits of customers over time.  `order_dow` is formatted in number so I created `order_dow_label` variable to easily notice the day of week. People usually buy their groceries on `r day_order$order_dow_label[1]` the most.


For example, in the first row of the dataset:

* `user_id`: `r instacart$user_id[1]`  
* `order_id`: `r instacart$order_id[1]`  
* `order_hour_of_day` : `r instacart$order_hour_of_day[1]`
* `product_name`: `r instacart$product_name[1]`  
* `aisle`: `r instacart$aisle[1]`


This shows that customer with id `r instacart$user_id[1]` placed  `r instacart$order_id[1]` order at `r instacart$order_hour_of_day[1]` o'clock and purchased `r instacart$product_name[1]` from the `r instacart$aisle[1]` aisle.


**How many aisles are there, and which aisles are the most items ordered from?**

```{r}
instacart |> 
  group_by(aisle) |> 
  summarise(count = n()) |> 
  arrange(desc(count))
```

The total number of unique aisles in the instacart dataset is `r length(unique(instacart$aisle_id))` and the most items are ordered from fresh vegetables aisle. 



**Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.**

```{r fig.height=8}
instacart |> 
  group_by(aisle) |> 
  summarise(count = n()) |> 
  filter(count > 10000) |> 
  ggplot(aes(x = reorder(aisle, count), y = count)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Number of Items Ordered by Aisle (>10,000)",
    x = "Aisle",
    y = "Number of Items Ordered"
  ) +
  theme_minimal(base_size = 13)
```



**Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.**

```{r}
instacart |> 
  filter(
    aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")
    ) |>
  group_by(aisle, product_name) |> 
  summarize(n_items = n(), .groups = "drop") |> 
  arrange(aisle, desc(n_items)) |> 
  slice_max(n = 3, order_by = n_items, by = aisle) |>
  knitr::kable()
```



**Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).**

As `order_dow` variable has values from 0 to 6, I converted these numbers into human-readable weekday names in variable `order_dow_label.`

```{r}
instacart |> 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |> 
  mutate(
    order_dow_label = wday(order_dow + 1, label = TRUE, abbr = FALSE)
  ) |> 
  group_by(product_name, order_dow_label) |> 
  summarize(
    mean_hour = round(mean(order_hour_of_day, na.rm = TRUE), digit = 2), 
    .groups = "drop"
  ) |> 
  pivot_wider(
    names_from = order_dow_label,
    values_from = mean_hour
  ) |> 
  knitr::kable()
```



# Problem 2

**Import, clean, and tidy Zillow datasets **
Import, clean, tidy NYC zipcode dataset 

```{r}
zipcode = 
  read_csv("data/Zip Codes.csv") |> 
  janitor::clean_names()
```

Import, clean, tidy  Zillow Observed Rent Index (ZORI) in New York City dataset

```{r}
zillow_nyc = 
  read_csv("data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv") |> 
  janitor::clean_names() |>
  rename(
    zip_code = region_name
  ) |> 
  pivot_longer(
    x2015_01_31:x2024_08_31,
    names_to = "date",
    values_to = "zori",
    names_prefix = "x"
  ) |> 
  mutate(
    county_name = str_remove(county_name, " County$")
  ) 
```



**There are 116 months between January 2015 and August 2024. How many ZIP codes are observed 116 times? How many are observed fewer than 10 times? Why are some ZIP codes are observed rarely and others observed in each month?**

(Question! Should I use the merged data of zipcode and zillow_nyc to consider the all the zipcodes instead of using only zillow dataset?)...!

```{r}
zipcode_all = 
  zillow_nyc |> 
  drop_na(zori) |> 
  group_by(zip_code) |> 
  summarize(
    n_month = n()
  ) |> 
  filter(n_month == 116)
```

```{r}
zipcode_10counts = 
  zillow_nyc |> 
  drop_na(zori) |> 
  group_by(zip_code) |> 
  summarize(
    n_month = n()
  ) |> 
  filter(n_month < 10)
```

`r nrow(zipcode_all)` zipcodes are observed 116 times and `r nrow(zipcode_10counts)` zipcodes are observed fewer than 10 times from January 2015 to August 2024. These zipcodes are rarely observed likely due to limited data or missing/incomplete data. For example, some ZIP codes have more housing units tracked by Zillow, others fewer. ZIP codes with very few units or newly created ZIP codes may appear only rarely. Additionally, reporting gaps can cause a month of data for certain ZIP codes to be missing.



**Create a table showing the average rental price in each borough and year (not month). Comment on trends in this table.**

```{r}
zillow_nyc |> 
  mutate(year = str_sub(date, 1, 4)) |> 
  group_by(county_name, year) |> 
  summarize(
    avg_rental_price = mean(zori, na.rm = TRUE)
  ) |> 
  knitr::kable(
    digits = 0,
    caption = "Average Rental Price (ZORI) by NYC counties and Year",
    col.names = c("Borough (County)", "Year", "Average ZORI")
  )
```

From the table, we can see the upward trend in rental prices over time for all boroughs in New York City. However, in 2021, both Kings and Queens has dropped compared to year of 2020. This was mainly due to the COVID-19 pandemic. By 2022 and 2023, prices in all boroughs, including Kings and Queens, rebounded sharply, surpassing pre-pandemic price. 



**Make a plot showing NYC Rental Prices within ZIP codes for all available years. Your plot should facilitate comparisons across boroughs. Comment on any significant elements of this plot.**


```{r}
rental_allyear = 
  zillow_nyc |> 
  mutate(year = as.integer(str_sub(date, 1, 4)))  |> 
  filter(!is.na(zori)) |>  
  group_by(zip_code, county_name, year) |> 
  summarize(avg_zori = mean(zori, na.rm = TRUE), .groups = "drop") |> 
  ggplot(aes(x = year, y = avg_zori, group = zip_code, color = county_name)) +
  geom_point() +
  facet_grid(. ~ county_name) +
  labs(
    title = "Rental Price Trends Across Boroughs from 2015 to 2024",
    x = "Year",
    y = "ZORI (Rental Price Index)"
  ) + 
  scale_x_continuous(
    breaks = c(2015, 2017, 2019, 2021, 2023),
    labels = c("2015", "2017", "2019", "2021", "2023")
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    legend.position = "bottom"
  ) 

rental_allyear
```

This plot shows rental price trends in New York City across ZIP codes and counties from 2015 to 2024. Only Richmond county has a record from 2020 to 2024. The Zillow Observed Rent Index (ZORI) has a steady upward trend, with a noticeable acceleration after 2021. The thin lines indicate individual ZIP codes, showing substantial within-county variation. New York county has the widest spread up until 8000 ZORI and it stands out with the highest rent levels among five counties and the steepest increase from 2021 to 2022. 



**Compute the average rental price within each ZIP code over each month in 2023. Make a reader-friendly plot showing the distribution of ZIP-code-level rental prices across boroughs; put differently, your plot should facilitate the comparison of the distribution of average rental prices across boroughs. Comment on this plot.**

!! question!
The question says over each month but the zillow data already provides the monthly based rental price for each zip code. Does it mean that we need to compute the average over 2023?
```{r}
rental_2023 = 
  zillow_nyc |> 
  separate(
    date, into = c("year", "month", "day")
  )  |> 
  filter(year == 2023) |> 
  group_by(zip_code, county_name) |> 
  summarize(avg_zori = mean(zori)) |> 
  ggplot(aes(x = county_name, y = avg_zori)) +
  geom_boxplot() +
  labs(
    title = "Rental Prices Across Boroughs in 2023",    
    x = "County",
    y = "Average ZORI (Rental Price Index)"
    
  ) +
  theme_minimal()

rental_2023
```

The plot shows the rental prices across counties in 2023. There are differences in median rental prices and the speadness of prices among zipcodes within each county. New York has the highest median rent price and the greatest variability having two outliers. On the other hands, Rickmond has the least variability and the lowest median rental prices. 

**Combine the two previous plots into a single graphic, and export this to a results folder in your repository.**

```{r}
rental_allyear / rental_2023 

rental_plot = rental_allyear + rental_2023 + plot_layout(ncol = 2)

ggsave("results/rental_combined.png", rental_plot, width = 12, height = 5, dpi = 300)

```



# Problem 3

**Load, tidy, merge, and organize the data sets.**
- exclude participants less than 21 years of age, and those with missing demographic data
- encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).

```{r}
#accelerometer data 
accel = 
  read_csv("data/nhanes_accel.csv") |> 
  janitor::clean_names() 

#demographic data
covar = 
  read_csv("data/nhanes_covar.csv", skip = 4) |> 
  janitor::clean_names() |> 
  filter(age >= 21) |> 
  drop_na() |> 
  mutate(
    sex = case_match(
      sex,
      1 ~ "male",
      2 ~ "female"
    ),
    education = case_match(
      education,
      1 ~ "Less than high school",
      2 ~ "High school equivalent",
      3 ~ "More than high school"
    )
  )
```

**Produce a reader-friendly table for the number of men and women in each education category**

```{r}
final_merged = 
  left_join(covar, accel, by = "seqn") 

final_merged |> 
  group_by(education, sex) |> 
  summarize(
    n_count = n(), 
    .groups = "drop"
  ) |> 
  knitr::kable() 
```

**create a visualization of the age distributions for men and women in each education category.**

```{r}
final_merged |> 
  ggplot(aes(x = education, y = age, fill = sex)) +
  geom_violin(trim = FALSE, alpha = .5) +
  labs(
    title = "Age Distributions by Education and Sex",
    x = "Education Level",
    y = "Age",
    fill = "Sex"
  ) +
  theme_minimal()
```

From the summarized table of the number of men and women in each education category, both sex with more than high school education represent the largest proportion of participants compared to the other education levels. The violin plot shows that individuals with more than high school education are concentrated in their 30s, while those with less than high school and high school equivalent education level groups show a broader and older age distribution. This demonstrates that higher education levels are common among young adults while old people are likely to have lower education level. 




**Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant.**

```{r}
total = 
  final_merged |> 
  mutate(
    total_activity = rowSums(select(final_merged, starts_with("min")), na.rm = TRUE))
```



**Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences.**

```{r}
total |> 
  ggplot(aes(x = age, y = total_activity, color = sex)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = 'loess', se = FALSE) +
  facet_grid(. ~ education) +
  labs(
    title = "Total Activity vs. Age by Education and Sex",
    x = "Age",
    y = "Total Activity (sum across minutes)",
    color = "Sex"
  ) +
  theme_minimal()
```

The plot illustrates the relationship between total activity and age, separated by sex and education level. There is a declining trend in total activity with increasing age, showing that older people tend to be less total activity than younger ones. The trend lines show peaks in activity during ages 30-40, followed by a steady decline afterward. Especially, participants with less than high school education level exhibit the steepest decline showing the largest difference between age groups. Women tend to have higher activity levels than man in the younger to middle-aged groups. Only in less than high school group, men surpass women in activity level after their 40s. Overall, both age and education level are important factors influencing activity levels. 



**Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.**


```{r, message=FALSE}
final_merged |>
  pivot_longer(
    cols = min1:min1440,
    names_to = "time",
    values_to = "activity"
  ) |>
  mutate(time = as.numeric(gsub("min", "", time))) |> 
  ggplot(aes(x = time, y = activity, color = sex)) +
  geom_smooth(se = FALSE, linewidth = 1) +
  facet_wrap(education ~ .) +
  labs(
    title = "24-hour Activity Pattern by Education Level and Sex",
    x = "time (minutes)",
    y = "Activity Level",
    color = "Sex"
  ) +
  theme_minimal() + 
  theme(legend.position = "bottom")
```

The plot displays the 24-hour activity level trends by education level and sex. Across all groups, activity levels are low during the early morning (around 0–300 minutes), rise sharply around mid-morning, and peak during daytime (approximately 600–1000 minutes), followed by a steady decline toward nighttime (after 1000 minutes).

Individuals with less than high school education show a sharper rise and the highest daytime peaks, possibly reflecting irregular daily routines. Across all education levels, men and women show broadly similar temporal activity trends. However, women tend to have slightly higher daytime activity levels across the day.